---
title: "tidyverse_dplyr_Advanced"
author: "Dr. Sampth Lonka"
date: "Created 9 January 2021; updated `r format(Sys.time(), '%B %d, %Y at %H:%M:%S')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# What is the tidyverse?

* tidyverse is a coherent collection of packages for data science
* Originally developed by Hadley Wickham (`ggplot2`,`plyr`)
* Packages include:
    * Data wrangling: `dplyr`, `tidyr`, `tibble`, `readr`
    * Visualization: `ggplot2`
    * List manipulation: `purrr`
    
# Why the tidyverse?
* Packages have a very consistent API
* Function names and commands follow a focused grammar
* Extremely powerful when working with data frames and lists (matrices, not so much, yet!)
* Allows pipes (`%>%` operator) to fluidly glue functionality together
* Very active developer and user community
* **Main advantage:** at its best, tidyverse data wrangling code can be read like a story using the pipe operator!
* Two keys to work :
   * learn about pipes `%>%`
   * learn the `dplyr` verbs
   
   
### The glorius pipe

* Tidyverse functions are at their best when composed together using the pipe operator
* It looks like this: `%>%`. Shortcut: use `ctrl + shift + m` in RStudio
* This operator actually comes from the `magrittr` package (automatically included in tidyverse)
* So it can be used on its own, completely independently of the `tidyverse`

```{r tidyverse}
library(tidyverse)
```

### Reading pip
Passing a single argument through pipes, we interpret something like:
```{r}
#x %>% f %>% g %>% h
```
as $h(g(f(x)))$
   
**Key takeaway: ** in your mind, when you see `%>%`, read this as **“and then”**

### Example

We can write `exp(1)` with pipes as `1 %>% exp()`, and `log(exp(1))` as `1 %>% exp() %>% log()`
```{r}
1 %>% exp()
```
```{r}
# computing log(exp(1))
1 %>% exp() %>% log()
```

### Read pipes for multiple arguments

Now for multi-arguments functions, we interpret something like:
```{r}
#x %>% f(y) 
```
as $f(x,y)$

### Load induilt data frame
```{r}
data <- mtcars # Renaming dataset
dim(data)
```

### Examples

```{r}
data %>% 
  head(4) # To read first 4 rows of the data frame
```

```{r results='hide'}
# base R
head(data, 4)
```
Notice that, with pipes:

 * Your code is more readable (arguably)
 * You can run partial commands more easily

# dplyr verbs
Our dplyr journey started learning with the following verbs (functions):

* filter(): subset rows based on logical criteria
* select(): select certain columns
* arrange(): order rows by value of a column
* mutate(): create new columns

In this section we will learn more verbs:

* slice(): subset rows based on integer indexing
* pull(): pull out a individual column
* rename(): rename columns
* mutate_at(): apply a function to given columns

 **Key takeaway:** think of data frames as nouns and dplyr verbs as actions that you apply to manipulate them - especially natural when using pipes

Use slice() when you want to indicate certain row numbers need to be kept:


```{r}
data %>% # inbuilt data frame
  slice(c(7,8,14:15)) # Reading rows 7,8, 14 and 15
```


```{r results='hide'}
# Base R
data[c(7,8,14:15),]
```

We can also do negative slicing:

```{r}
data %>% 
  slice(-c(1:2,19:23)) %>% 
  nrow() # count the remaining the rows in the data frame
```
```{r results='hide'}
# Base R:
nrow(data[-c(1:2,19:23),])
```

### `pull()`

* You can grab a single column from a data frame and get it back as a vector if you use pull
* select preserves column structure even with a single column

```{r}
data %>% pull(mpg)
```


```{r}
# Same as: data$mpg

data %>% select(mpg)
```


### `mutate()`

Use `mutate()` when you want to create one or several columns:

```{r results='hide'}
data <- data %>% 
  mutate(hp_wt = hp/wt, 
         mpg_wt = mpg/wt) 

# Base R:
#data$hp_wt <- data$hp/mtcars$wt
#data$mpg_wt <- data$mpg/mtcars$wt
```

```{r}
data <- data %>% 
  mutate(hp_wt_correct = hp/wt,
         hp_wt_cyl = hp_wt_correct/cyl) 
data %>% head(2)
```

```{r}
data <- data %>% 
  mutate(hp_wt = 1) # update hp_wt to just the one value
data %>% head(2)
```

### `mutate_at()`

Use `mutate_at()` when you want to apply a function to one or several columns:

```{r}
# correction
data <- data %>% mutate(hp_wt = hp_wt_correct)

data <- data %>%
  mutate_at(c("hp_wt", "mpg_wt"), log) 
```

```{r}
# Base R:
#data$hp_wt <- log(data$hp_wt)
#data$mpg_wt <- log(data$mpg_wt)
```

### `rename()`

Use `rename()` to easily rename columns:
```{r}
data %>% 
  rename(hp_wt_log = hp_wt, mpg_wt_log = mpg_wt) %>%
  head(2)
```

### Note

Calling `dplyr` verbs always outputs a new data frame, it does not alter the existing data frame

# Join operations

A “join” operation in database terminology is a merging of two data frames for us. There are 4 types of joins:

* **Inner join** (or **just join**): retain just the rows each table that match the condition
* **Left outer join** (or **just left join**): retain all rows in the first table, and just the rows in the second table that match the condition
* **Right outer join** (or **just right join**): retain just the rows in the first table that match the condition, and all rows in the second table
* **Full outer join** (or **just full join**): retain all rows in both tables
* Column values that cannot be filled in are assigned NA values

### Two toy data frames
```{r}
has_kids_tab1 <- data.frame(name = c("Robert Downey, Jr", "Scarlett Johansson", "Chris Hemsworth"), 
                  children = c(3, 1, 3),
                stringsAsFactors = FALSE)
# `stringsAsFactors` indicates whether strings in a data frame should be treated as factor variables or as just plain strings.
americans_tab2 <- data.frame(name = c("Chris Evans", "Robert Downey, Jr", "Scarlett Johansson"),
                  age = c(38, 54, 34),
                  stringsAsFactors = FALSE)
has_kids_tab1

# With base R functions, to avoid conversion of strings to factors you would do, for example:
# 
# x = read.csv("my_file.csv", stringsAsFactors=FALSE)
# In readr you can just read the file, as there is no stringsAsFactors argument and no automatic conversion of strings to factors:
# 
# library(readr)
# x = read_csv("my_file.csv")
```

```{r}
americans_tab2
```

## inner_join()
Suppose we want to join tab1 and tab2 by name, but keep only actors in intersection (aka in both tables):

```{r}
inner_join(x = has_kids_tab1, y = americans_tab2, by = "name")
```

Suppose we want to join tab1 and tab2 by name, but keep all actors from tab1:


```{r}
left_join(x = has_kids_tab1, y = americans_tab2, by = c("name" = "name"))
```

## right_join()

Suppose we want to join tab1 and tab2 by name, but keep all actors from tab2:

```{r}
right_join(x = has_kids_tab1, y = americans_tab2, by = "name")
```

## full_join()

Finally, suppose we want to join tab1 and tab2 by name, and keep all actors from both:

```{r}
full_join(x = has_kids_tab1, y = americans_tab2, by = "name")
```




# Summary

* The `tidyverse` is a collection of packages for common data science tasks
* Tidyverse functionality is greatly enhanced using pipes (`%>%` operator)
* Pipes allow you to string together commands to get a flow of results
* `dplyr` is a package for data wrangling, with several key verbs (functions)
* `slice()` and `filter()`: subset rows based on numbers or conditions
* `select()` and `pull`: select columns or pull out as single column vector
* `arrange()`: order rows by one or multiple columns
* `rename()` and `mutate()`: rename or create columns
* `mutate_at()`: apply a function to given columns
* `*_join()` where * = inner, left, right, or full: join two data frames together according to common values in certain columns, and `*` indicates how many rows to keep



# References

* [tidyverse.org](https://www.tidyverse.org/): the main place for package updates and news
* [tidyverse forum](https://community.rstudio.com/c/tidyverse/6): the official forum for questions
* [dplyr cheatsheet (current version) from RStudio](https://rstudio.com/resources/cheatsheets/#dplyr)

