---
title: "Exercise-PCA"
author: "Bala"
date: "22/03/2021"
output: word_document
editor_options: 
  chunk_output_type: inline
---

Answer the following:

Compute the sample covariance matrix S
Compute eigen values and eigen vectors of S
Perform PCA
What is your suggested first k− principal components. Explain through scree plot.
Perform graphical analysis of k components.
Can you make any inference using principle components.

```{r}
#loading libraries
library(tidyverse)
library(knitr)
library(broom)
library(ggfortify)
```
```{r}
#loading dataset
swiss %>% head(10) %>% knitr::kable()
DATA <- swiss
```
```{r}
# Sample Covariance matrix
covmatrix <-  cov(DATA)
covmatrix
```
```{r}
#Eigen values and eigen vectors
eigvals <- eigen(covmatrix)
eigvals$values
sum(eigvals$values)
eigvals$vectors[,1]
```
```{r}
#PCA
comp <- prcomp(DATA)
summary(comp)
eigvals$vectors
names(DATA)
# Here PC1 covers 74.6 , PC2 18.1 and remaining constituting to 7.3
```
```{r}
#Scree plot
x <- c(1:6)
y <- eigvals$values
plot(x,y,type="l",main="scree plot")
# Since there is the elbow bend at 2 , the suggested first k pricipal components are 2 i.e. 2
```
```{r}
# Components
PC <- as.matrix(DATA) %*% eigvals$vectors
ggplot(PC,aes(PC[,1],PC[,2]))+
  geom_point()
```

#Analysis from principal components
PC1 and PC2 cumulative proportions captures up to 0.92 data and adding PC3 results in 0.98.





  