---
title: "MiniProject"
author: "Bala"
date: "09/03/2021"
output:
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: inline
---
#columns
```{r}
library(corrplot)
library(caret)
library(lattice)
library(car)
library(tidyverse)
library(MASS)
```

1.age: The person’s age in years
2.sex: The person’s sex (0 = female, 1 = male)
3.cp: The chest pain experienced (Value 1: typical angina, Value 2: atypical angina, Value 3: non-anginal pain, Value 4: asymptomatic)
trestbps: The person’s resting blood pressure (mm Hg on admission to the hospital)
4.chol: The person’s cholesterol measurement in mg/dl
5.fbs: The person’s fasting blood sugar (if > 120 mg/dl, 1 = true; 0 = false)
6.restecg: Resting electrocardiographic measurement (0 = normal, 1 = having ST-T wave abnormality, 2 = showing probable or definite left ventricular hypertrophy by Estes’ criteria)
7.thalach: The person’s maximum heart rate achieved
exang: Exercise induced angina (1 = yes; 0 = no)
8.oldpeak: ST depression induced by exercise relative to rest (‘ST’ relates to positions on the ECG plot)
9.slope: the slope of the peak exercise ST segment (Value 1: upsloping, Value 2: flat, Value 3: downsloping)
10.ca: The number of major vessels (0-3)
11.thal: A blood disorder called thalassemia (1 = normal; 2 = fixed defect; 3 = reversable defect)
12.target: Heart disease (0 = no, 1 = yes)

- Basic Statistics (you may include summary of statistics, correlation matrix, ...)
- EDA (Histograms, Scatterplots, Boxplots,....)
- Linear Models (Choose your dependent and independent variables wisely)
- ANOVA (Choose your classes and variable to perform ANOVA)
- Logistic regression

Basic Statistics and EDA
-------
```{r}
heartdata1<- read.csv("/home/nagendra/Downloads/heart.csv")
head(heartdata1)
heartdata<- as_tibble(heartdata1)

#Looking at the data columns and their importance with respect to the dependent variable target 
```
```{r}
str(heartdata)
```

```{r}
summary(heartdata)
#Age here varied from 29 to 77 with 54 as average age
#here average of gender is 0.6832 signifying most heart patients are of female category
#chest pain varied from 0 to 3 level with 1 as average signifying only less patients have high chest pain levels
#Blood pressure while resting varied from 94
```

```{r}

corr<-cor(heartdata)
corrplot(corr,method= "number",type= "full")
# Here oldpeak and slope are negatively 58 % correlated and all other variables are not significantly correlated.
```

```{r}

featurePlot(x=heartdata[,c("age","sex","cp","trestbps","chol","fbs","restecg","target")],y=heartdata$thalach)
```

```{r}
ggplot(heartdata, aes(x=thalach)) + 
    geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.5,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666") 
```

```{r}
# Bar plot for target (yes or no)
heartdata$target <- as.factor(heartdata$target)
ggplot(heartdata, aes(x=heartdata$target, fill=heartdata$target)) + 
  geom_bar() +
  xlab("Heart Disease") +
  ylab("Count") +
  ggtitle("Analysis of Presence and Absence of Heart Disease") +
  scale_fill_discrete(name = "Heart Disease", labels = c("Absence", "Presence"))

```

```{r}
# Group the different ages in three groups (young, middle aged, old)
young <- heartdata[which((heartdata$age<45)), ]
middleaged <- heartdata[which((heartdata$age>=45)&(heartdata$age<55)), ]
elderly <- heartdata[which(heartdata$age>55), ]
groups <- data.frame(age_group = c("young","middle aged","elderly"), group_count = c(NROW(young$age), NROW(middleaged$age), NROW(elderly$age)))

#ploting different age groups
ggplot(groups, aes(x=groups$age_group, y=groups$group_count, fill=groups$age_group)) + 
  ggtitle("Age Analysis") +
  xlab("Age Group")  +
  ylab("group Count") +
  geom_bar(stat="identity") +
  scale_fill_discrete(name = "Age Group", labels = c("Elderly", "Middle Aged", "Young"))
```

```{r}
# Adding the age groups to the dataset
heartdata <- cbind(heartdata, groups = ifelse((heartdata$age<45), 0, ifelse((heartdata$age>=45)&(heartdata$age<55), 1, 2)))
heartdata$groups <- as.factor(heartdata$groups)
# we will remove the age column as this is very generalised column and we have divided it, group, to include 
# that in our analysis more specifically.

heartdata = subset(heartdata, select = c(-age))
# Discrete vs Discrete vs Discrete variable: age_group ~ target ~ sex
ggplot(heartdata, aes(x= factor(heartdata$groups), y=heartdata$sex, colour=target)) + 
  geom_boxplot(stat = "boxplot",
               position = "dodge2") +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(width = 0.2) +
  xlab("Age Groups") +
  ylab("Gender") +
  ggtitle("Analysis of gender with different age group with presence or absense of heart disease")
```

```{r}
# Bar plot for sex
ggplot(heartdata, aes(x= sex, fill=target)) + 
  geom_bar() +
  xlab("Gender") +
  ylab("Gender Count") +
  ggtitle("Analysis of Gender") +
  scale_fill_discrete(name = "Heart disease", labels = c("No", "Yes"))
```

```{r}
#cp (Chest Pain) variable Analysis
# Bar plot for The chest pain experienced 
ggplot(heartdata, aes(x= cp, fill=cp)) + 
  geom_bar() +
  xlab("Chest Pain Type") +
  ylab("Count") +
  ggtitle("Analysis of Chest Pain Experienced") +
  scale_fill_discrete(name = "Chest Pain Type", labels = c("Typical angina pain", "Atypical angina pain", "Non-Anginal pain", "Asymptomatic pain"))

```

```{r}
# Bar plot for The chest pain ~ target
ggplot(heartdata, aes(x= cp, fill=target)) + 
  geom_bar() +
  xlab("Chest Pain Type") +
  ylab("Count") +
  ggtitle("Analysis of Chest Pain Experienced") +
  scale_fill_discrete(name = "Heart disease", labels = c("No", "Yes"))
```

```{r}
# Density graph for trestbps (resting blood pressure)
ggplot(heartdata, aes(x = trestbps, fill = target)) +
  geom_density(alpha=0.5) +
  scale_fill_discrete(name = "Heart disease", labels = c("No", "Yes"))
```

```{r}
# Density plot for oldpeak ~ target
ggplot(heartdata, aes(x = oldpeak, fill = target)) +
  geom_density(alpha=0.5) +
  xlab("ST depression induced") +
  ylab("Count") +
  ggtitle("Analysis of ST depression induced and presence of heart disease") +
  scale_fill_discrete(name = "Heart disease", labels = c("No", "Yes"))
```

```{r}
# Density plot for thalach ~ target
ggplot(heartdata, aes(x = thalach, fill = target)) +
  geom_density(alpha=0.5) +
  xlab("Maximum Heart Rate Achieved") +
  ylab("Count") +
  ggtitle("Analysis of relation of heart rate with presence of heart disease") +
  scale_fill_discrete(name = "Heart disease", labels = c("No", "Yes"))

```

```{r}
par(mfrow= c(2,4))
#boxplot(heartdata$age, main="Age",col = "red")
boxplot(heartdata$chol, main="Cholestrol",col = "yellow")
boxplot(heartdata$thalach, main="Thalach",col="green")
boxplot(heartdata$slope,main="Slope",col = "blue")
#
```

```{r}
tab= table(heartdata$sex)
pie(tab,main="pie chart")

#More Males in the population compared to females in the dataset
```

```{r}
#Scatterplot
plot <- ggplot(heartdata, aes(trestbps, chol))
plot + geom_point(aes(colour = factor(cp)))
```

```{r}

mod1<- lm(thalach~.,data=heartdata)
par(mfrow=c(2,4))
plot(mod1)
summary(mod1)

#Here our model is yielding very less Adjusted r square implying that it is not normal
#Let's try another model
```

```{r}
mod2<- lm(thalach ~ exang + slope + target + cp + trestbps + chol,data=heartdata)
par(mfrow=c(2,4))
plot(mod2)
summary(mod2)
#Here adjusted r square has increased but not significantly
#We will go for another
```
```{r}
mod3<- lm(thalach ~ groups * groups + exang*exang + slope * slope + target + cp * cp + chol + trestbps,data = heartdata)
par(mfrow=c(2,4))
plot(mod2)
summary(mod3)
#Here we yielded the same adj r square, looks like we reached the saturation point...
```
```{r}
head(heartdata)
```



```{r}
library(broom)
summary_stats <- tbl_df(bind_rows(glance(mod1) %>% dplyr::select(adj.r.squared,sigma,AIC,BIC),
glance(mod2) %>% dplyr::select(adj.r.squared,sigma,AIC,BIC),
glance(mod3) %>% dplyr::select(adj.r.squared,sigma,AIC,BIC)))
summary_stats 
#so this is the summary of the stats
```

```{r}
bc<-boxcox(thalach~groups+cp+trestbps+chol+exang+slope+target
,data=heartdata)
lambda <- bc$x[which.max(bc$y)]
lambda

heartdata$y <- ((heartdata$thalach)^lambda-1/lambda)
Newmodel <- lm(thalach~groups+cp+trestbps+chol+exang+slope+target,data=heartdata)
summary(Newmodel)
shapiro.test(Newmodel$residuals)

glance(Newmodel) %>% dplyr::select(adj.r.squared,sigma,AIC,BIC)
#Even after applying boxcox we clearly see that there is no improvement in the adj r square 
```

```{r}
data= sample_n(heartdata,10)#Taking some random samples from b_data
predict(Newmodel,data)#Predicted values
data$thalach#Actual values
```

```{r}
step.model <- stepAIC(mod1,direction = "backward",trace = TRUE)
summary(step.model)
```

ANOVA
-------

```{r}
heartdata$cd=0
 heartdata$cd[heartdata$cp==0] ="Zero"
 heartdata$cd[heartdata$cp==1] ="One"
 heartdata$cd[heartdata$cp==2] ="Two"
 heartdata$cd[heartdata$cp==3] ="Three"
 

```
```{r}
changer <- function(type){p_data =heartdata%>%filter(cp==type)
return(sample_n(p_data,20))}

```
```{r}
t_zero = changer(0)
t_one = changer(1)
t_two = changer(2)
t_three=changer(3)
adata = bind_rows(t_zero,t_one,t_two,t_three)
dim(adata)

```

```{r}
res_aov <- aov(thalach~ cd,data = adata)
summary(res_aov)
shapiro.test(res_aov$residuals)
leveneTest(thalach~ cd, data=adata)
#variances are not equal. Accept alternate hypothesis.
```

```{r}
#$H_0:$ Variances are equal
#$H_1:$ at least one varaince is different
Final<- TukeyHSD(res_aov)
Final
plot(Final)
```


Logistic Regression
--------


```{r}
#We can have many threshold to classify the problem.We will find optimal using a package and use it
#To check whether the Exercise Induced Angia is present or not using number of floroscopy induced arteries and another...
```

```{r}
dim(heartdata)
```

```{r}
# To have same output
set.seed(123)
# Indexing the sample data
default_idx = sample(nrow(heartdata), 240)
default_trn = heartdata[default_idx, ]
default_tst = heartdata[-default_idx, ]
dim(default_trn)
```

```{r}
#Model using age ,trestbps,cholestorol , thalach , and oldpeak...

model1 = glm(exang~groups+trestbps+chol+thalach+oldpeak,data=default_trn,family='binomial')
model_n_pred = ifelse(predict(model1, type = "link") > 0.5, "1", "0")
table_trn = table(predicted = model_n_pred, actual = default_trn$exang)
predicted = predict(model1, default_trn, type="response")
confusion_matrix = confusionMatrix(table_trn, positive = "1")
c(confusion_matrix$overall["Accuracy"], 
  confusion_matrix$byClass["Sensitivity"], 
  confusion_matrix$byClass["Specificity"])
#Here sensitivity is not at all significant which shows that the prediction of true positive is low...
#let's go for next model
```

```{r}
model2 = glm(exang~groups+thalach+oldpeak+target,data=default_trn,family='binomial')
model_n_pred = ifelse(predict(model2, type = "link") > 0.5, "1", "0")
table_trn = table(predicted = model_n_pred, actual = default_trn$exang)
predicted = predict(model2, default_trn, type="response")
confusion_matrix2 = confusionMatrix(table_trn, positive = "1")
c(confusion_matrix2$overall["Accuracy"], 
  confusion_matrix2$byClass["Sensitivity"], 
  confusion_matrix2$byClass["Specificity"])
#Improvement is apparent but not substantial
#And hunt goes on for another

```

```{r}
model3=glm(exang~groups+sex+cp+trestbps+chol+fbs+restecg+thalach+oldpeak+slope+ca+thal+target
,data=default_trn,family='binomial')
car::vif(model3)
model_n_pred = ifelse(predict(model3, type = "link") > 0.001, "1", "0")
table_trn = table(predicted = model_n_pred, actual = default_trn$exang)
predicted = predict(model3, default_trn, type="response")
confusion_matrix3 = confusionMatrix(table_trn, positive = "1")
c(confusion_matrix3$overall["Accuracy"], 
  confusion_matrix3$byClass["Sensitivity"], 
  confusion_matrix3$byClass["Specificity"])
#And here we go with so far the best model yielding the sensitivity at .66 and specificity at .89
#So this is the best logistic model for now.
```


