---
title: "MDSC-206-Assignment-IV-20226"
author: "Bala"
date: "25/02/2021"
output: word_document
editor_options: 
  chunk_output_type: inline
---
Assignment-IV:
===
This assignment for 20 Marks

 **Answer the following for the dataset 'ToothGrowth' from CRAN**
 
- Perform Exploratory Data Analysis (must include scatter plots, box-plots) (5 Marks)
- Choose your qualitative and quantitative variables in given data and perform One-Way ANOVA (5 Marks)
- Choose your variables from the data to perform Two-Way ANOVA (5 Marks)
- Presentation (5 Marks)
  
```{r}
#Loading the required libraries
library(tidyverse)
library(car)
library(lattice)
library(ggplot2)
```
```{r}
head(ToothGrowth)
glimpse(ToothGrowth)
```
There are 60 Observations and 3 features. We will take two features from the data namely `len` and `dose`


```{r}
mydata<- ToothGrowth 
head(mydata)
```
The data with all the three features...



```{r}
summary(mydata)
```
```{r}
#Scatterplot
ggplot(mydata)+aes(dose,len,color=supp)+geom_point()+labs(title = "Len vs Dose")
```
For 0.5 mg/ml dosage, Length of the teeth varied between each guinea pig from 1 to 21,Not 
For 1.0 mg/ml dosage, Length of the teeth varied between each guinea pig
from 12 to 27,
For 2.0 mg/ml dosage, Length of the teeth varied between each guinea pig
from 15 to 35...
All these are not exactly conclusive as there are many outliers....
In conclusion, The more the dosage higher the growth of length.
```{r}
#Boxplot
par(mfrow=c(1,3))
boxplot(mydata$len,main="Length of the Teeth",col = "red")
```
The black line shows the mean of the data, and there are many outliers in length...




```{r}
mydata$supp<- as.factor(mydata$supp)
mydata$dose<- as.factor(mydata$dose)

```
Converting the doses to factors as we just have 3 classes 0.5 , 1 , 2...





One Way ANOVA
===
```{r}
data1 <- ToothGrowth %>% 
            select(len,supp)
head(data1)
```
The null and alternative hypothesis of an ANOVA are:
- $H_0:$ The effect of the two supplements effect on length of the teeth is equal(OJ growth = VC growth)
- $H_1:$ Their effect on the teeth is not equal

Residuals following normal distribution
Equality of Variance 

```{r}
res_aov <- aov(len~supp,data=data1)
summary(res_aov)
```

 Given that the p-value is greater than 0.05, we accept the null hypothesis, so we accept the hypothesis that the effect of two supplements on length of the teeth is equal

```{r}
par(mfrow=c(1,2))
plot(res_aov,2)
plot(res_aov,3)
```

```{r}
hist(res_aov$residuals)

qqnorm(res_aov$residuals,xlab = "norm quantiles")
```

```{r}
shapiro.test(res_aov$residuals)
```
 Either by observing plots or `shapiro.test()` results, we can conclude that the residuals are normal. 

Testing Hypothesis for Levense Test:

- $H_0:$ Variances are equal
- $H_1:$ at least one variance is different
```{r}
leveneTest(len~supp,data=data1)
```
The p-value being larger than the significance level of 0.05, we do not reject the null hypothesis, so we cannot reject the hypothesis that variances are equal between supplements (p-value = 0.275).
```{r}
res_aov2 <- aov(len~supp*dose,data=mydata)
summary(res_aov2)
```
As p-values are near to zero for two factors, we reject the null hypothesis. Thus there is a significance difference between in the popuplation mean across the factors.
```{r}
hist(res_aov2$residuals)
```

```{r}
shapiro.test(res_aov2$residuals)
```
By looking at the pvalue we can say that it is normal
```{r}
leveneTest(len~supp*dose,data = mydata)
```
pvalue is greater than 0.05 therefore variances are equal...
```{r}
par(mfrow=c(1,2))
plot(res_aov2,2)
plot(res_aov2,3)
```

Conclusions:
-> The higher the dosage the higher the growth of length of teeth...
-> The effect of two supplements is nearly same for above 2 mg/ml cases.




